---
title: "Exploring The Essen and Densmore"
author: "David John Baker"
date: "April 30, 2018"
output:
  pdf_document: default
  html_document: default
---

## Import Data

First thing needed is to import all of the data.

```{r}
library(data.table)
library(ggplot2)
densmore <- fread("../data/feature_data/densmoreFeatures.csv") # Native American Songs
essen <- fread("../data/feature_data/essenFeatures.csv") # Essen Collection
```

## Explore Data 

Check out all the variables listed.

```{r}
names(essen)
```

Special subset of this data has region of Germany associated with it. 

```{r}
unique(essen$Region)
unique(densmore$Country)
# Link to what key of each one is can be found here http://kern.humdrum.org/cgi-bin/browse?l=essen/europa/deutschl
```

### Explore Large Level Features of Corpus 

Swap in other variables below!

```{r}
ggplot(essen, aes(glob.duration)) + geom_density() + labs(title = "Density Plot of Length")

ggplot(essen, aes(mean.entropy)) + geom_density() + labs(title = "Density Plot of Mean Entropy")
```


### Relationship Between Variables

You can use ```ggplot2``` to explore the relationship between different variables.
Try to swap out the x and y variables with others on a continuous scale.
You can also group by categorical variables.

```{r}
ggplot(essen, 
       aes(x = len, y = p.entropy))+ geom_point() + labs(main = "Length of Melody vs Pitch Entropy")


ggplot(essen, 
       aes(x = len, y = tonalness))+ geom_point() + labs(main = "Length of Melody vs Tonalness")

ggplot(essen, 
       aes(x = len, y = tonalness, color = mode))+ geom_point() + labs(main = "Length of Melody vs Tonalness, By Mode")


ggplot(essen, 
       aes(x = len, y = tonalness, color = Country))+ geom_point() + labs(main = "Length of Melody vs Tonalness, By Type")

```


### Subsetting

You can "zoom" in on some variables using R's indexing functions.

```{r}
# Index within the plot 
ggplot(essen[glob.duration > 7 & glob.duration < 9], 
       aes(x = len, y = tonalness, color = mode))+ geom_point() + labs(main = "Length of Melody vs Tonalness, By Mode")

# Try different group by 
ggplot(essen[glob.duration > 7 & glob.duration < 9], 
       aes(x = len, y = tonalness, color = Country))+ geom_point() + labs(main = "Length of Melody vs Tonalness, By Country")

# Convert to Z Scale
ggplot(essen[glob.duration > 7 & glob.duration < 9], 
       aes(x = scale(len), y = scale(tonalness), color = Country))+ geom_point() + labs(main = "Length of Melody vs Tonalness, By Country")

# Convert to Z Scale
ggplot(densmore, 
       aes(x = note.dens, y = tonalness, color = Country))+ geom_point() + labs(main = "Densmore Density vs Tonalness, By Tribe")

# Look at Group Differences
ggplot(densmore,
       aes(y = tonalness, x = Country)) + geom_boxplot() + labs(title = "Differences in Tonalness between Native American Tribes")

ggplot(essen,
       aes(y = tonalness, x = mode)) + geom_boxplot() + labs("Differences in Tonalness in Essen Collection Split by Mode")

```




